|

### 2021-08-09 appendix/appendix.Rmd
latex changes

```{r, include=FALSE}
## Starting
XXXX

# Required packages
XXXX

# Load datasets
XXXX
```

@@

1. The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
2. The mayor proposes a policy vector $(r, g, \bf{x}, \bf{t})_M$.
3. The city council votes the proposal.
  - If the council accepts, then the policy is implemented, and the game ends. 
  - Otherwise, the reversal policy is implemented.
4. (Reversal Policy) At the beginning of each iteration of the reversal policy, one councilor is randomly recognized to be the proposer. She proposes a vector $(\bf{x}, \bf{t})$, with $r =  g = 0$. If the proposal is accepted, it is implemented. If the proposal is rejected, a new round of reversal start, but with the municipal resources reduced by a factor $\delta$.

@@

However, the decrease in rents is not sufficient to have lower rent extraction. The political costs, if high enough, can offset the gains with lower rents. Higher political costs result from having more members of the opposition in the legislature. The next proposition shows the sufficient condition for having a lower cost of govern when we increase the number of legislators.

There exists an ex-ante probability of electing a councilor aligned with the government $\overline{\gamma}$ such that, for all $\gamma > \overline{\gamma}$, the cost of govern decreases as $N$ increases, i.e. $C_G(N+1) \leq C_G(N)$.

@@

If the proposition is correct, then $C_G(N) - C_G(N+1) \geq 0$. Let us rewrite the results as:

\[
\dfrac{N}{2}(\underline{t} + \underline{x}_o^g(N-1)) + \sum_{\eta_g = 0}^{N/2}\mathbb{P}(\eta_g)\left( \underline{x}_m^g(N-1) - \underline{x}_m^o(N-1) \right) - \left(\dfrac{N}{2}+1\right)(\underline{t} + \underline{x}_o^g(N+1)) - \sum_{\eta_g = 0}^{N/2+1}\mathbb{P}(\eta_g)\left( \underline{x}_m^g(N+1) - \underline{x}_m^o(N+1) \right) \geq 0
\]

@@

@@

@@

\subsection{Variable Sources and Descriptive Statistics\label{appsources}}

We use three information sources, either from Brazilian governmental agencies or available online. Table \ref{varsources} displays the primary sources and their respective URLs.

\begin{sidewaystable}[!htbp] \centering 
  \caption{Data Sources} 
  \label{varsources}
  \scalebox{0.75}{
\begin{tabular}{@{\extracolsep{5pt}}llll} 
\\[-1.8ex]\hline\hline \\[-1.8ex] 
Source Code & \multicolumn{1}{c}{Source Name} & \multicolumn{1}{c}{Description} & \multicolumn{1}{c}{URL} \\ 
\hline \\[-1.8ex] 
& & & \\
DataSUS & Brazilian Health Ministry Data Service & Collects data on Health Care & \url{http://www.datasus.gov.br} \\
& & & \\
IBGE & Brazilian Institute of Geography and Statistics & Collects data Geography, Economics and Demography & \url{http://www.ibge.gov.br} \\ 
& & & \\
INEP & Data Service of Ministry of Education & Collects data on performance of Education & \url{http://www.inep.gov.br} \\ 
& & & \\
InterLegis & Senate Legislative Data Service & Collect Data on Legislative in Brazil & \url{http://www.interlegis.leg.br} \\
& & & \\
IPEA & Brazilian Institute of Applied Economics & Collects data on the Economy & \url{http://www.ipeadata.gov.br} \\ 
& & & \\
MDS & Social Security Ministry & Collect data on Social Coverage and Effectiveness of Social Programs & \url{http://www.mds.gov.br} \\ 
& & & \\
TSE & Superior Electoral Tribunal & Collects data on elections results & \url{http://www.tse.jus.br} \\
& & & \\
\hline\hline \\[-1.8ex] 
\end{tabular} 
}
\end{sidewaystable}

Now let us define each variable.

\textbf{Outcomes Aggregated at the Municipal Level:}

\begin{itemize}
    \item \textbf{Number of Seats 2000}: Number of councilors in the municipality by the 2000 elections (source: TSE).
    \item \textbf{Population 2000}: Municipal population according to the 2000 Brazilian Census (source: IBGE).
    \item \textbf{Per-Capita GDP Census 2000 (in millions)}: Municipal per-capita GDP computed by the IBGE Census (source: IPEA).
    \item \textbf{Proportion of Poverty Census 2000}: Proportion of poverty in the municipality defined as people living on less than R\$ 70.00 a day (source: MDS).
    \item \textbf{Number of Seats 2004}: Number of city-councilors in a municipality according to the Electoral Justice decision (source: TSE).
    \item \textbf{Infant Mortality 2005--2008}: Infant mortality computed as the number of children born alive that died before reaching one year of age divided by the number of children born alive, multiplied by 1,000 (source: DataSUS)
    \item \textbf{Post-Natal Mortality Rate 2005--2008}: Infant mortality computed as the number of children born alive that successfully lived 28 days, but died before reaching one year of age, divided by the number of children born alive, multiplied by 1,000 (source: DataSUS)
    \item \textbf{Coverage Family Health Program 2005--2008}: Number of families covered by the Family Health Program (source: DataSUS).
    \item \textbf{Prop. Born w. 7+ Pre-Natal 2005--2008}: Proportion of children born who went to seven or more prenatal care consultations (source: DataSUS)
    \item \textbf{Enrollment Elementary School 2005--2008}: Number of children enrolled in elementary schools, averaged by classroom size (K--4) (source: INEP).
    \item \textbf{Enrollment Middle School 2005--2008}: Number of children enrolled in middle schools, averaged by classroom size (5--9) (source: INEP)
    \item \textbf{Quality of Elementary School Index 2005--2008}: IDEB score averaged by schools in the municipality. IDEB scores are composed by the student's grades in math and language, multiplied by an indicator of the distortion between the year the child is supposed to be and the year the child is. The estimators are cleaned to avoid influences of schools and classroom specific effects (source: INEP, years: 2005, 2007).
    \item \textbf{Quality of Middle School Index 2005--2008}: IDEB score averaged by schools in the municipality. IDEB scores are composed by the student's grade in math and language, multiplied by a score capturing the distortion between the year the child is supposed to be in, and the year the child is studying. The estimators are cleaned to avoid the influence of schools and classrooms specific effects. (source: INEP, years: 2005, 2007).
    \item \textbf{Mayoral Pre-Electoral Coalition Size 2004}: Number of elected councilors in the mayoral pre-electoral coalition (source: TSE)
    \item \textbf{Number of Appointed Bureaucrats 2005--2008}: Number of bureaucrats of the direct administration that were appointed to jobs in the municipality (source: IBGE, years: 2005, 2006, and 2008)
    \item \textbf{Number of Career Bureaucrats 2005--2008}: Number of career bureaucrats in the municipality. Career bureaucrats cannot be fired by politicians and earn considerably better pensions after retirement (source: IBGE, years: 2005, 2006, and 2008)
    \item \textbf{Number of Councilor's Appointed Assistants 2005}: Number of assistants appointed for the councilor cabinet (source: InterLegis)
    \item \textbf{Number of Females Elected 2004}: Number of females elected to city council (source: TSE)
    \item \textbf{Number of Non-Whites Elected 2004}: Number of non-whites elected to city council. We collected this data for municipalities less than 10,000 inhabitants away from the cutoffs (source: our compilation based on the TSE candidate pictures). 
    \item \textbf{Competition per Seat 2004}: Number of people running for city-councilor divided by the number of city councilors in the municipality (source: TSE)
    \item \textbf{Proportion Approved Legislation 2005}: Number of approved legislation in 2005 divided by the number of proposed legislation (source: InterLegis)
\end{itemize}

\textbf{Legislation Dataset:}

We coded the legislation approved by the councilors in 64 municipalities that are 10 thousand inhabitants away from the council size thresholds. We classified the legislation in five categories:

\begin{itemize}
    \item \textbf{Local Public Goods}: Legislation that provides a service targeted at the individual, group, or geographical level.
    \item \textbf{Public Goods or General Legislation}: Legislation that provides a service for the entire municipality, or approves a law that does not discriminate among citizens.
    \item \textbf{Oversight}: Legislation that requests information to the mayor's office or the bureaucracy about the provision of services.
    \item \textbf{Others}: Legislation that is not classified as Local Public Goods, Public Goods, or Oversight. Usually honors or procedures.
    \item \textbf{Education and Health Care}: Legislation about Education or Health Care provision.
\end{itemize}

\textbf{Online 2016 Former City Councilors Survey:}

We run an online survey asking 174 councilors whether helping with [...service...] gives [none, few, some, or many] votes. The services were:

\begin{itemize}
    \item \textbf{Help Hospital Admission}: We asked former councilors whether helping with hospital admission gives [none, few, some, many] votes (source: survey).
    \item \textbf{Help School Admission}: We asked former councilors whether helping with school admission gives [none, few, some, many] votes (source: survey).
    \item \textbf{Help Getting Medication}: We asked former councilors whether helping voters get medication gives [none, few, some, many] votes (source: survey).
    \item \textbf{Fix Street Potholes}: We asked former councilors whether pressing the bureaucracy to fix street potholes gives [none, few, some, many] votes (source: survey).
    \item \textbf{Other Local Public Goods}: We asked former councilors whether helping voters with local public goods and services such as traffic rules, sewage constructions, potable water, and others, gives [none, few, some, many] votes (source: survey).
    \item \textbf{Legislative Duties}: We asked former councilors whether the act to discuss, propose, and vote on legislation gives [none, few, some, many] votes (source: survey).
    \item \textbf{Honors Legislation}: We asked former councilors whether proposing bills to praise local celebrities gives [none, few, some, many] votes (source: survey).
    \item \textbf{Oversee School Quality}: We asked former councilors whether overseeing school quality gives [none, few, some, many] votes (source: survey).
    \item \textbf{Oversee Construction}: We asked former councilors whether overseeing public construction gives [none, few, some, many] votes (source: survey)
\end{itemize}

In this Appendix, we present other questions we asked in the survey. They are meant to improve the knowledge we have about the city councilors perceptions about their representation. We have not added these to the main paper because they were more complimentary and not of general interest. The descriptive statistics of the variables used in the main paper follow in Table \ref{tabdescrvarspaper}.

\input{tabs/edited/sumstatOut2016e.tex}

\subsection{Threshold Manipulation, Sorting, and Pre-treatment Consistency}

The validity of our causal claim relies on three assumptions. First, the impossibility of municipalities to manipulate the placement at the left or right-hand side of the thresholds. Second, that we are consistently comparing municipalities in the same cutoffs. Lastly, that we have no pre-treatment differences in our sample.

For the first assumption, \citet{McCrary2008} proposes a measure of the distributional imbalance around the discontinuity, testing whether cases are more abundant in the left or the right of the cutoff. For our research, the McCrary statistic is 0.391 (SE $=$ 0.299), showing no evidence of manipulation.

We run the \citet{cattaneo2019simple} manipulation test, which is based on the density of the local polynomial estimator. We use local polynomial orders from one to four. As the null hypothesis is no manipulation, the p-values for each polynomial order are: local linear (p-value $=0.442$); quadratic (p-value $=0.740$); cubic (p-value $=0.998$); and quartic (p-value $=0.620$). Therefore, there is no manipulation according to \citet{calonico2019regression}.

We also graph the \citet{McCrary2008} test, pooling all the discontinuities. The results follow in Figure \ref{manipulationplot} and show no signs of manipulation.

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=5in,angle=0]{figs/McCrary.pdf}}
      \caption{\citet{McCrary2008} manipulation test}
      \bigskip
      \label{manipulationplot}
    \end{minipage}
  \end{center}
\end{figure}

For the second assumption, we know that running RDD in multiple thresholds may lead to inconsistent estimates when there is an imbalance in the running variable's distribution around the cutoffs \citep{Bertanha2015, Cattaneo2016, eggers2018regression}. For example, in our dataset, we have twelve discontinuities, and at each threshold, we have a sharp increase of one councilor, from 9 to 21. When running an RDD on these discontinuities, a consistent estimator should fit an exact change of one councilor. However, if we pool all the discontinuities together with no correction, we find an increase in council size of 1.63 councilors. This imbalance is because when we combine all the discontinuities, we are implying that changes in municipalities right below the 47,619 cutoffs (9 to 10 councilors) are comparable to changes in towns right above the 571,428 cutoffs (20 to 21 council members). That is incorrect, not only because we could be comparing a municipality with nine councilors with one with 21 councilors, but also because these municipalities diverge in many other aspects, such as population to GDP per capita.

To avoid this problem, we show by simulation that adding controls, especially the variables responsible for the multiple threshold assignments (in our case, the population in 2003), improves the consistency and efficiency of the estimates. Relying on this fact, we add five controls to our estimates: population in 2003; GDP per capita; the number of seats in 2000; year; and a dummy for the northeast municipalities. The reason for the first variable is the multiple threshold assignments. We add GDP per capita because wealthier cities tend to be more productive and have better public services. The number of seats in 2000 intends to control for the fact that some municipalities could experience a change provoked by the previous council size, confounding our estimates. The year and the northeast dummies serve to improve efficiency. Adding controls improves efficiency in RDD \citep{calonico2019regression} and also enhances the consistency in the multiple thresholds RDD. See the following section for further tests and simulations.

For the last assumption, we should have no changes in pre-treatment covariates. These pre-treatment covariates are variables measured before the primary outcomes. As the threshold rule selected by the Brazilian Electoral Court was unpredictable, it should not detect any variation before the thresholds were in place.

Table \ref{validitycheck} displays the results of the validity checks. At the top of the table, we run the first-stage regressions, both using and not using the controlling covariates strategy proposed, for both real and the placebo cutoffs. At the bottom, we estimate the pre-treatment covariate balance.

\begin{table}[htb!] \centering 
  \caption{Research Design Validity Check} 
  \label{validitycheck} 
    \scalebox{0.9}{
  \begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
 & \multicolumn{4}{c}{\textbf{Panel A: Validity Check --- First Stage}} \\ \cline{2-5}
   & \multicolumn{1}{c}{\small{(1)}} & 
 \multicolumn{1}{c}{\small{(2)}} & 
 \multicolumn{1}{c}{\small{(3)}} & 
 \multicolumn{1}{c}{\small{(4)}} \\

 & \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} & 
 \multicolumn{1}{c}{\small{Additional Num.}} & 
 \multicolumn{1}{c}{\small{Placebo Add.}} \\ 
  & \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} & 
  \multicolumn{1}{c}{\small{Seats 2004}} & 
  \multicolumn{1}{c}{\small{Num. Seats 2004}} \\
  & \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(Without Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} & 
  \multicolumn{1}{c}{\small{(With Controls)}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
\\[-1.8ex] 
 LATE & 1.63$^{***}$ & -0.31$^{**}$ & 1.00$^{***}$ & -0.00 \\ 
  & (0.51) & (0.14) & (0.0004) & (0.0000) \\ 
\hdashline \\[-1.8ex]   
 N Left & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{4621} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{906} \\ 
Eff N Left & \multicolumn{1}{c}{199} & \multicolumn{1}{c}{638} & \multicolumn{1}{c}{49} & \multicolumn{1}{c}{886} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{385} & \multicolumn{1}{c}{51} & \multicolumn{1}{c}{477} \\ 
BW Loc Poly & 8.717 & 6.700 & 3.008 & 8.629 \\ 
BW Bias & 13.620 & 12.114 & 5.008 & 13.466 \\ 
\midrule 
 & \multicolumn{4}{c}{\textbf{Panel B: Validity Check --- Pre-Treatment Variables}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(5)}} & 
 \multicolumn{1}{c}{\small{(6)}} & 
 \multicolumn{1}{c}{\small{(7)}} & 
 \multicolumn{1}{c}{\small{(8)}} \\
  & \multicolumn{1}{c}{\small{Number of}} & 
 \multicolumn{1}{c}{\small{Population}} & 
 \multicolumn{1}{c}{\small{GDP}} & 
 \multicolumn{1}{c}{\small{\% of Poverty}} \\ 
  & \multicolumn{1}{c}{\small{Seats 2000}} & 
  \multicolumn{1}{c}{\small{2000 Census}} & 
  \multicolumn{1}{c}{\small{2000 Census}} &
   \multicolumn{1}{c}{\small{2000 Census}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
 LATE & 0.17 & -2057.60 & 0.06 & -9.72 \\ 
  & (0.67) & (1926.33) & (0.18) & (6.27) \\ 
  \hdashline \\[-1.8ex]
 N Left & \multicolumn{1}{c}{5178} & \multicolumn{1}{c}{5131} & \multicolumn{1}{c}{5131} & \multicolumn{1}{c}{5131} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{343} \\ 
Eff N Left & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{219} & \multicolumn{1}{c}{200} & \multicolumn{1}{c}{194} \\ 
Eff N Right & \multicolumn{1}{c}{147} & \multicolumn{1}{c}{157} & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{142} \\ 
BW Loc Poly & 8.970 & 9.368 & 8.739 & 8.530 \\ 
BW Bias & 13.860 & 15.029 & 14.752 & 13.035 \\ 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $ ^{***} $ p $ < $ .01; $ ^{**} $ p $ < $ .05; $ ^{*} $ p $ < $ .1. RD local linear estimates using \citet{calonico2019regression} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population, GDP per capita, number of seats in 2000, year, and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observation in the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

Notice that without controls, the first stage overestimates the change in the primary treatment and underestimates the effect in the placebo regressions. After adding covariates, the results improve, and the point estimates are now one for the primary model and zero for the placebo regressions, which are the correct values. Moreover, Panel B of Table \ref{validitycheck} shows that the pre-treatment covariates remain unchanged around the thresholds. These tests reinforce the validity of our research designs.

\pagebreak

\subsection{Population Thresholds in Brazil Before and After the 2003 Supreme Court Decision\label{appidentif}}

The identification strategy requires exogenous changes in the city council size thresholds. Previous to the changes we exploit here, the 1988 Brazilian Constitution states the following distribution of seats per population size, presented in Table \ref{tabcitycouncilsize1988}.

\begin{table}[htp!]
\centering
\caption{City council size from the 1988 Constitution}
\scalebox{0.75}{
\begin{tabular}{rccrrr}
  \hline\hline
 & \multicolumn{1}{c}{Min. Leg.} & \multicolumn{1}{c}{Max. Leg.} & \multicolumn{1}{c}{Min. Pop.} & \multicolumn{1}{c}{Max. Pop.} & \multicolumn{1}{c}{Num. Mun. Bin} \\ 
 & & & & & \multicolumn{1}{c}{(2001-2004)} \\
  \hline
  1 & 9  & 21 & 0 & 1,000,000 &  5,537 \\ 
  2 & 33 & 41 & 1,000,001 & 5,000,000 & 11 \\ 
  3 & 42 & 55 & 5,000,001 & $\infty$ & 2 \\ 
   \hline\hline
\end{tabular}
}
\label{tabcitycouncilsize1988}
\end{table}

These cutoffs were in place until 2003, when the municipality of Mira Estrela decided to decrease the council size from eleven to nine councilors. The two losing councilors started a judicial contestation that made it to the Supreme Court. The court ruled that the Constitution meant that council size should vary proportionally with population, starting with nine legislators, until the maximum number of councilors, 55. Following this decision, the Superior Electoral Tribunal (TSE) decided the thresholds that should be followed in the 2004 election (Resolution 21,702/2004). According to their decision, the legislature should increase as displayed in Table \ref{tabcitycouncilsize2004all}.

\begin{table}[htp!]
\centering
\caption{City council size thresholds according to TSE resolution 21,702/2004}
\scalebox{0.75}{
\begin{tabular}{rcrrr}
  \hline\hline
 &  \multicolumn{1}{c}{Num. Leg.} & \multicolumn{1}{c}{Min. Pop.} & \multicolumn{1}{c}{Max. Pop.} & \multicolumn{1}{c}{Num. Mun. Bin} \\
 & & & & \multicolumn{1}{c}{(2003 pop. proj.)} \\
  \hline
  1 &  9 & 0 & 47,619 & 5,029 \\ 
  2 &  10 & 47,620 & 95,238 & 317 \\ 
  3 &  11 & 95,239 & 142,857 & 89 \\ 
  4 &  12 & 142,858 & 190,476 & 43 \\ 
  5 &  13 & 190,477 & 238,095 & 30 \\ 
  6 &  14 & 238,096 & 285,714 & 21 \\ 
  7 &  15 & 285,715 & 333,333 & 13 \\ 
  8 &  16 & 333,334 & 380,952 & 13 \\ 
  9 &  17 & 380,953 & 428,571 & 6 \\ 
  10  & 18 & 428,572 & 476,190 & 6 \\ 
  11  & 19 & 476,191 & 523,809 & 4 \\ 
  12  & 20 & 523,810 & 571,428 & 5 \\ 
  13  & 21 & 571,429 & 1,000,000 & 13 \\ 
  14  & 33 & 1,000,001 & 1,121,952 & 1 \\ 
  15  & 34 & 1,121,953 & 1,243,903 & 2 \\ 
  16  & 35 & 1,243,904 & 1,365,854 & 1 \\ 
  17  & 36 & 1,365,855 & 1,487,805 & 2 \\ 
  18  & 37 & 1,487,806 & 1,609,756 & 1 \\ 
  19  & 38 & 1,609,757 & 1,731,707 & 1 \\ 
  20  & 39 & 1,731,708 & 1,853,658 & 0 \\ 
  21  & 40 & 1,853,659 & 1,975,609 & 0 \\ 
  22  & 41 & 1,975,610 & 4,999,999 & 4 \\ 
  23  & 42 & 5,000,000 & 5,119,047 & 0 \\ 
  24  & 43 & 5,119,048 & 5,238,094 & 0 \\ 
  25  & 44 & 5,238,095 & 5,357,141 & 0 \\ 
  26  & 45 & 5,357,142 & 5,476,188 & 0 \\ 
  27  & 46 & 5,476,189 & 5,595,235 & 0 \\ 
  28  & 47 & 5,595,236 & 5,714,282 & 0 \\ 
  29  & 48 & 5,714,283 & 5,833,329 & 0 \\ 
  30  & 49 & 5,833,330 & 5,952,376 & 0 \\ 
  31  & 50 & 5,952,377 & 6,071,423 & 1 \\ 
  32  & 51 & 6,071,424 & 6,190,470 & 0 \\ 
  33  & 52 & 6,190,471 & 6,309,517 & 0 \\ 
  34  & 53 & 6,309,518 & 6,428,564 & 0 \\ 
  35  & 54 & 6,428,565 & 6,547,611 & 1 \\ 
  36  & 55 & 6,547,612 & $\infty$  & 0 \\ 
   \hline\hline
\end{tabular}
}
\label{tabcitycouncilsize2004all}
\end{table}

As there are few municipalities above one million inhabitants, we decided to restrict our attention to municipalities with less than $571,429 + 47,619 = 619,048$ inhabitants. Table \ref{tabcitycouncilsize2004short} displays the frequency of municipality by thresholds used in our paper.

\begin{table}[htp!]
\centering
\caption{City council thresholds used in the paper}
\scalebox{0.85}{
\renewcommand*{\arraystretch}{1.2}
\begin{tabular}{rcrrr}
  \hline \hline
 & \multicolumn{1}{c}{Legislature} & \multicolumn{1}{c}{Min.} & \multicolumn{1}{c}{Max.} & \multicolumn{1}{c}{Num. Mun. Bin} \\
 & \multicolumn{1}{c}{Size} & \multicolumn{1}{c}{Population} & \multicolumn{1}{c}{Population} & \multicolumn{1}{c}{(2003 pop.)} \\
  \hline
  1 & 9 & 0 & 47,619 & 5,029 \\ 
  2 & 10 & 47,620 & 95,238 & 317 \\ 
  3 & 11 & 95,239 & 142,857 & 89 \\ 
  4 & 12 & 142,858 & 190,476 & 43 \\ 
  5 & 13 & 190,477 & 238,095 & 30 \\ 
  6 & 14 & 238,096 & 285,714 & 21 \\ 
  7 & 15 & 285,715 & 333,333 & 13 \\ 
  8 & 16 & 333,334 & 380,952 & 13 \\ 
  9 & 17 & 380,953 & 428,571 & 6 \\ 
  10 & 18 & 428,572 & 476,190 & 6 \\ 
  11 & 19 & 476,191 & 523,809 & 4 \\ 
  12 & 20 & 523,810 & 571,428 & 5 \\ 
  13 & 21 & 571,429 & 1,000,000 & 13 \\ 
   \hline \hline
\end{tabular}
}
\label{tabcitycouncilsize2004short}
\end{table}

The Supreme Court ruled that the electoral courts should use the 2003 population projection to assign municipalities to cutoffs. In Brazil, the Brazilian Institute of Geography and Statistics (IBGE) propose the official population projections. Sorting is not an issue for our identification strategy, as IBGE is an insulated bureaucracy, insulated from local influences (see also the McCrary test above).

To provide an overview of the data dispersion, Figure \ref{sample_plot} plots the municipalities by their proximity from the cutoffs. In the map we provide the contour of every Brazilian municipality, coloring it by the proximity to the council size thresholds. The darker the color, the closer to the population thresholds. In regression discontinuity, the closer a municipality is to the cutoff, the more influential it is in the estimation. Figure \ref{sample_plot} also shows that our municipalities are reasonably distributed around the whole country.

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=7in,angle=0]{figs/sampplot2019.pdf}}
      \caption{Municipalities by their Selection Status}
      \bigskip
      \label{sample_plot}
    \end{minipage}
  \end{center}
\end{figure}

\pagebreak

Moreover, we could question whether the discontinuities studied here are sharp or fuzzy. In order to be fuzzy, the laws passed by the Supreme Court had to be contested or not be perfectly enforced. For instance, \citet{cavalcanti2017essays} shows in his second chapter that there has been a misuse of these discontinuities. However, this problem was absent in our RDD, as the problem identified by Cavalvanti was based on the use of thresholds approved in 2009. In our case, we use the thresholds approved in 2003, and until a period before 2009 (in our case, 2008).

\subsection{Identification Strategy in Multiple Thresholds Regression Discontinuity Design\label{ISMT}}

Usually when carrying out regression discontinuity with multiple cutoffs, authors pool all the discontinuities together and estimate the effect of council size on outcomes. Let $C = \{c_0, c_1, \cdots, c_n, c_{n+1}\}$ be the set of cutoffs thresholds associated with the running variable $X$. The outcome variable is assumed to be $Y$. In the pooled model, the Local Average Treatment Effect is:

\[
LATE_{nocontrols} \quad = \quad \E \left[ \lim_{x \downarrow c_i} \E[Y_i | X_i] - \lim_{x \uparrow c_i} \E[Y_i | X_i] | \forall c_i \in C \right]
\]

Nevertheless, this may be problematic as the municipal characteristics in each cutoff may vary from cutoff to cutoff. The population imbalance certainly carries out heterogeneity, in the way shown by \citet{Cattaneo2016}. To ensure we have comparability, and that we estimate the average LATE for all cutoffs, avoiding most of the heterogeneity, we propose to add a set of controls denoted by $Z_i$ that vary between cutoffs. For instance, the population that determines the change from one threshold to another also determines all the cutoffs. As the population varies smoothly around each cutoff, it makes it a perfect control to add. The new estimator is:

\[
LATE_{controls} \quad = \quad \E \left[ \lim_{x \downarrow c_i} \E[Y_i | X_i] - \lim_{x \uparrow c_i} \E[Y_i | X_i] | \forall c_i \in C, Z_i \right]
\]

We control for Population, GDP, a dummy for the Northeast States, seats before the 2003 decision, and year. All these control variables are smooth in each cutoff, but are significantly different from cutoff to cutoff, making them adequate control variables for our case \citep{calonico2019regression}. To perform the estimations, we use a triangular kernel, which places more weight on the municipalities that are closer to each cutoff. To compute the optimal bandwidth, we use the \citet{CCT2014} method. To study the sensitivity to the bandwidth choice, we vary the bandwidth from 50\% to 200\% of the optimal bandwidth size. We also use cluster-robust standard errors at the municipal level.

However, as it is uncommon for a regression discontinuity to use control variables, some readers may ask why in this case is it advisable? To answer this question, we run a series of simulations to show that controlling for the variable that determines the assignment not only helps improve the consistency of our estimates but also improves their efficiency.

Consider the data distributed in one of the four forms depicted by Figure \ref{simRDdd}. We add nine thresholds, at 0.1, 0.2, and so on until 0.9, and we create six types of outcomes. The first three outcomes are sharp changes. In the first, we add variation from one to 10, with one step for each change. In the second, we add 1 in the first cut, zero in the second cut, and -1 in the third cut, repeating this pattern for the remaining cutoffs. The estimated change should be equal to zero. In the third model, we add one in the first cut, $1+0.9$ in the second cut, $1+0.9+0.8$ in the third cut, and so on until the last cutoff. This is intended to simulate a diminishing effect from one cutoff to the other. The last three outcomes are the same as depicted here, added by a Normal random disturbance, mean zero, and variance 0.01. We then run a thousand Monte Carlo simulation for each combination, $4 \times 6 \times 2 = 48$ in total, fitting a regression with and without the running variable as a control.

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{0.9\linewidth}
      \centerline{\includegraphics[width=6in,angle=0]{figs/simdistrRDD.pdf}}
      \caption{Simulation RD with Multiple Thresholds -- Data Distribution}
      \bigskip
      \label{simRDdd}
    \end{minipage}
  \end{center}
\end{figure}

\pagebreak

A consistent estimator should fit one in the first and fourth data distribution.  In the second and fifth, the change should be of $(1, 0, -1, 1, 0, -1, 1, 0, -1)$, and the average change here is equal to zero. In the third and sixth models the change should be equal to $(0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1)$, and the average change should be equal to $0.5$. The first three models should present a more efficient estimation than the last three models. We display the results in Figure \ref{estimDGP}, showing that the controlled models are more consistent and efficient.

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{1\linewidth}
      \centerline{\includegraphics[width=4.5in,angle=0]{figs/estimsRDD.pdf}}
      \caption{Estimations for the Different Data Generating Processes}
      \bigskip
      \label{estimDGP}
    \end{minipage}
  \end{center}
\end{figure}

\subsection{Placebo Regressions for the Aggregated Mechanism Outcomes\label{placmechmain}}

>>>>

```{r, include=FALSE}
## Starting
#XXXX

# Required packages
#XXXX

# Load datasets
#XXXX
```

@@

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, \bf{x}, \bf{t})_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts, then the policy is implemented, and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) At the beginning of each iteration of the reversal policy, one councilor is randomly recognized to be the proposer. She proposes a vector $(\bf{x}, \bf{t})$, with $r =  g = 0$. If the proposal is accepted, it is implemented. If the proposal is rejected, a new round of reversal start, but with the municipal resources reduced by a factor $\delta$.
\end{enumerate}

@@

However, the decrease in rents is not sufficient to have lower costs to govern when we increase the number of legislators. The political costs, if high enough, can offset the gains with lower rents. Higher political costs result from having more members of the opposition in the legislature. The next proposition shows the sufficient condition for having a lower cost of govern when we increase the number of legislators.

\begin{proposition}\label{mainprop}

There exists an ex-ante probability of electing a councilor aligned with the government $\overline{\gamma}$ such that, for all $\gamma > \overline{\gamma}$, the cost of govern decreases as $N$ increases, i.e., $C_G(N+1) \leq C_G(N)$.

@@

Let us rewrite the results as:

\[
XXXX \geq 0
\]

@@

Let $\Delta\eta_n(\gamma) \equiv \left[\sum_{\eta_g = 0}^{N/2+1}\mathbb{P}(\eta_g)\eta_g - \sum_{\eta_g = 0}^{N/2}\mathbb{P}(\eta_g)\eta_g\right]$. There is no closed form solution for this equation. We can approximate the boundaries using @wXXXX but graphically, we can have a good sense of the XXXX:

XXXX GRAPH DELTA GAMMA HERE

And solving for $\Delta\eta_n(\gamma)$ we have that:

\[
XXXX
\]

The natural boundaries for $\Delta\eta_n(\gamma)$ are minus-one and one. Minus-one means that the additional legislature seat went to the opposition while one means that the additional legislature seat went to the government. XXXX

@@

Taking stocks, the model provides two testable hypothesis. Let a legislature that previously had $N$ legislators, and now increased the legislature by $N+1$. Then:

\begin{enumerate}
  \item[H1. ] The provision of public goods increase when the costs to govern decrease with legislature size.
  \item[H2. ] The costs to govern decrease with legislature size when the chance of electing a government aligned legislator is sufficiently high.
\end{enumerate}

## Paper Version of the Model

In the main paper, we simplify the analysis with two assumptions regarding the costs of govern. First, we assume a simpler reversal where there is only one reversal stage and after that a reversal policy of zero in all policy components take place. Second, we assume that the cost for govern will depend on the ex-ante expected number of legislator belonging to the government and the opposition. In this case, the rents $b(N)$ become:

\[
XXXX
\]

And Proposition \ref{mainprop} becomes:

XXXX

\[
XXXX
\]

\begin{proof}
XXXX
\end{proof}

All the dynamics remain the same but the second hypothesis becomes clearer, with a simplified quantity for the probability of electing a government-aligned politician, $\overline{\gamma}$. The simplified $\overline{\gamma}$ is helpful because it shows that the political concerns require a probability larger than one-half, with a precise amount depending on the XXXX.

@@
